---
title: "Mojave_Div_ITS_phyloseq"
author: "Nat Pombubpa"
date: "Updated on August 5, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###STEP1: Load all necessary packages for analysis
More information about Phyloseq can be found at the following link: [Phyloseq](https://joey711.github.io/phyloseq/)
If you get error in this step, you probably need to install any packages which causes error.

```{r warning=FALSE, message=FALSE}
library(ape)
library(vegan)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(plyr)
library(data.table)
library(tidyr)
library(tidyverse)
library(multcompView)
```

####Function to summarize taxonomy 
plot_taxa_summary(GlobalPatterns, "Phylum")
plot_taxa_summary(GlobalPatterns, "Phylum", "SampleType")
summarize_taxa(GlobalPatterns, "Phylum")
summarize_taxa(GlobalPatterns, "Phylum", "SampleType")

This function was written by Joey (but it is not part of phyloseq package)
[https://github.com/joey711/phyloseq/issues/418](https://github.com/joey711/phyloseq/issues/418)

```{r warning=FALSE}
fast_melt = function(physeq){
  # supports "naked" otu_table as `physeq` input.
  otutab = as(otu_table(physeq), "matrix")
  if(!taxa_are_rows(physeq)){otutab <- t(otutab)}
  otudt = data.table(otutab, keep.rownames = TRUE)
  setnames(otudt, "rn", "taxaID")
  # Enforce character taxaID key
  otudt[, taxaIDchar := as.character(taxaID)]
  otudt[, taxaID := NULL]
  setnames(otudt, "taxaIDchar", "taxaID")
  # Melt count table
  mdt = melt.data.table(otudt, 
                        id.vars = "taxaID",
                        variable.name = "SampleID",
                        value.name = "count")
  # Remove zeroes, NAs
  mdt <- mdt[count > 0][!is.na(count)]
  # Calculate relative abundance
  mdt[, RelativeAbundance := count / sum(count), by = SampleID]
  if(!is.null(tax_table(physeq, errorIfNULL = FALSE))){
    # If there is a tax_table, join with it. Otherwise, skip this join.
    taxdt = data.table(as(tax_table(physeq, errorIfNULL = TRUE), "matrix"), keep.rownames = TRUE)
    setnames(taxdt, "rn", "taxaID")
    # Enforce character taxaID key
    taxdt[, taxaIDchar := as.character(taxaID)]
    taxdt[, taxaID := NULL]
    setnames(taxdt, "taxaIDchar", "taxaID")
    # Join with tax table
    setkey(taxdt, "taxaID")
    setkey(mdt, "taxaID")
    mdt <- taxdt[mdt]
  }
  return(mdt)
}

summarize_taxa = function(physeq, Rank, GroupBy = NULL){
  Rank <- Rank[1]
  if(!Rank %in% rank_names(physeq)){
    message("The argument to `Rank` was:\n", Rank,
            "\nBut it was not found among taxonomic ranks:\n",
            paste0(rank_names(physeq), collapse = ", "), "\n",
            "Please check the list shown above and try again.")
  }
  if(!is.null(GroupBy)){
    GroupBy <- GroupBy[1]
    if(!GroupBy %in% sample_variables(physeq)){
      message("The argument to `GroupBy` was:\n", GroupBy,
              "\nBut it was not found among sample variables:\n",
              paste0(sample_variables(physeq), collapse = ", "), "\n",
              "Please check the list shown above and try again.")
    }
  }
  # Start with fast melt
  mdt = fast_melt(physeq)
  if(!is.null(GroupBy)){
    # Add the variable indicated in `GroupBy`, if provided.
    sdt = data.table(SampleID = sample_names(physeq),
                     var1 = get_variable(physeq, GroupBy))
    setnames(sdt, "var1", GroupBy)
    # Join
    setkey(sdt, SampleID)
    setkey(mdt, SampleID)
    mdt <- sdt[mdt]
  }
  # Summarize
  Nsamples = nsamples(physeq)
  summarydt = mdt[, list(meanRA = sum(RelativeAbundance)/Nsamples,
                         sdRA = sd(RelativeAbundance),
                         minRA = min(RelativeAbundance),
                         maxRA = max(RelativeAbundance)),
                  by = c(Rank, GroupBy)]
  return(summarydt)
}

plot_taxa_summary = function(physeq, Rank, GroupBy = NULL){
  # Get taxa summary table 
  dt1 = summarize_taxa(physeq, Rank = Rank, GroupBy = GroupBy)
  # Set factor appropriately for plotting
  RankCol = which(colnames(dt1) == Rank)
  setorder(dt1, -meanRA)
  dt1[, RankFac := factor(dt1[[Rank]], 
                          levels = rev(dt1[[Rank]]))]
  dt1[, ebarMax := max(c(0, min(meanRA + sdRA))), by = eval(Rank)]
  dt1[, ebarMin := max(c(0, min(meanRA - sdRA))), by = eval(Rank)]
  # Set zeroes to one-tenth the smallest value
  ebarMinFloor = dt1[(ebarMin > 0), min(ebarMin)]
  ebarMinFloor <- ebarMinFloor / 10
  dt1[(ebarMin == 0), ebarMin := ebarMinFloor]

  pRank = ggplot(dt1, aes(x = meanRA, y = RankFac)) +
    scale_x_log10() +
    xlab("Mean Relative Abundance") +
    ylab(Rank) +
    theme_bw()
  if(!is.null(GroupBy)){
    # pRank <- pRank + facet_wrap(facets = as.formula(paste("~", GroupBy)))
    pRank <- pRank + geom_point(mapping = aes_string(colour = GroupBy),
                                size = 5)
  } else {
    # Don't include error bars for faceted version
    pRank <- pRank + geom_errorbarh(aes(xmax = ebarMax,
                                        xmin = ebarMin))
  }
  return(pRank)
}
```


###STEP2: Import Mapping file (metadate file)
1.Check mapping file before import to R, R doesn’t seem to like sample name to start with number or contain “-” in sample name. If you get error in this step, you should check file name first.

2.First column of first row should not start with #, R will not read the first row that starts with #

3. You can choose which samples to include in analysis by indicating "KEEP" in Description column

```{r}
meta = read.table("Mojave_mappingfile_8-Aug-2018.txt",header=TRUE,row.names=1,sep="\t",stringsAsFactors=FALSE)
```

If need to keep or remove some samples
```{r}
#meta <- meta[which(meta$Layer %in% c("Surface")),]
```

###STEP3: Check if your metadata file has been import successfully and correctly

The output will show a table of your metadata file (mapping file).

*If you do not have header, you might start your first row with #

```{r warning=FALSE}
head(meta)
```

###STEP4: Construct sample_data-class using imported metadata

```{r}
sampleData <- sample_data(meta)
```

###STEP5: Import OTU table

OTU table from Mojave 16S data is “MC2017FC.otu_table.txt”.
MC-Mojave Crust-2017-FC-Fungi completed

```{r}
otus <- read.table("MC2018FC.otu_table.txt",header=T,sep="\t",row.names=1)
otumat <- as(as.matrix(otus), "matrix")
OTU = otu_table(otumat, taxa_are_rows = TRUE)
```

Check imported OTU table

```{r warning=FALSE}
head(OTU)
```

###STEP6: Import taxonomy table
Taxonmy table generated from AMPtk need to be rearranged using following script.

“perl rdp_taxonmy2mat.pl<Input_taxonmy.txt>Output_taxonomy.txt”

rdp_taxonomy2mat.pl was created by Professor Jason E. Stajich

```{r}
taxmat <- read.table("MC2018FC.taxonomy.fix.txt", header=T,sep="\t",row.names=1)
taxmat <- as(as.matrix(taxmat),"matrix")
TAX = tax_table(taxmat)
```

###STEP7: Import phylogenetic tree
Phylogenetic tree can also be include for further phylogenetic analysis.

```{r warning=FALSE}
treefile = "MC2018FC.tree.phy"
tree = read.tree(treefile)
```

###STEP8: Construct Phyloseq object
To construct phyloseq object, otu table, taxonomy table, and sampleData are required. Phylogenetic tree can be included, but it is not necessary for constructing phyloseq object.
Construct Phyloseq object called "Physeq"

```{r warning=FALSE}
physeq = phyloseq(OTU,TAX,sampleData,tree)
```

Check phyloseq object
This should indicate that your physeq is a "phyloseq-class experiment-level object""

```{r warning=FALSE}
physeq
```

###STEP9: Remove singletons
Remove any OTUs that present only one time.

```{r }
physeq.prune = prune_taxa(taxa_sums(physeq) > 1, physeq)
```

```{r warning=FALSE}
physeq.prune
```

###STEP10: Plot read counts to check dataset
Check read counts: any samples that have very low reads should be removed.
[Ref](http://evomics.org/wp-content/uploads/2016/01/phyloseq-Lab-01-Answers.html)

```{r}
readcount = data.table(as(sample_data(physeq.prune), "data.frame"),
                 TotalReads = sample_sums(physeq.prune), 
                 keep.rownames = TRUE)
setnames(readcount, "rn", "SampleID")

#For plotting, use command below.
#SeqDepth = ggplot(readcount, aes(TotalReads)) + geom_histogram() + ggtitle("Sequencing Depth")
```

TotalReads of all the samples can be in this table (select only SampleID and TotalReads columns).
In order to check samples with low number of reads, "order()" can be used to sort "TotalReads" column.
In this dataset, N55.Rhizo has very low number of reads, so will will filter this sample out using the next minimum number of reads.
```{r}
readcount = readcount[order(readcount$TotalReads), c("SampleID", "TotalReads")]
```

```{r}
head(readcount)
```

```{r warning=FALSE}
set.seed(1)
physeq.prune.rarefy = rarefy_even_depth(physeq.prune, sample.size = 6842, replace = FALSE, trimOTUs = FALSE)
physeq.prune.rarefy
```

###STEP11: Rarefaction curve
Extract OTU table, transpose and convert to data frame
```{bash}
#create new folder for figures if needed
#mkdir Figures
```

Extract OTU table from phyloseq object as data frame and use vegan to calculate rarefaction curve
```{r}
otu.rare <- otu_table(physeq.prune.rarefy)
otu.rare <- as.data.frame(t(otu.rare))
sample_names <- rownames(otu.rare)

#vegan rarecurve 
otu.rarecurve <- rarecurve(otu.rare, step = 500, sample = 30000, label = T)
```

```{r}
rare <- lapply(otu.rarecurve, function(x){
  b <- as.data.frame(x)
  b <- data.frame(OTU = b[,1], Sample.size = rownames(b))
  b$Sample.size <- as.numeric(gsub("N", "",  b$Sample.size))
  return(b)
})
```
label list
```{r}
names(rare) <- sample_names
```
convert to data frame
```{r}
#tidyverse
rare <- map_dfr(rare, function(x){
  z <- data.frame(x)
  return(z)
}, .id = "sample")
```

```{r}
head(rare)
```

Plot rarecurve using ggplot

```{r}
ggplot(data = rare)+
  geom_line(aes(x = Sample.size, y = OTU, color = sample), show.legend = FALSE)+
  scale_x_continuous(labels =  waiver()) + theme_bw() + ggtitle("Fungal Rarefaction Curve")+
  theme(plot.title = element_text(hjust = 0.5))
```

Save ggplot as pdf to figures folder using the full path

```{r}
pdf("./Figures/Fungal Rarefaction Curve.pdf", width = 8, height = 5 )
ggplot(data = rare)+
  geom_line(aes(x = Sample.size, y = OTU, color = sample), show.legend = FALSE)+
  scale_x_continuous(labels =  waiver()) + theme_bw() + ggtitle("Fungal Rarefaction Curve")+
  theme(plot.title = element_text(hjust = 0.5))
dev.off()
```

###STEP11.1: Plot Alpha diversity by Crust_type
Alpha diversity can be Chao1, Observed, Shannon, Simpson
This plot include statistical analysis using "stat_compare_means" with "method = anova"

```{r}
physeq.prune.rarefy.plot.richness.Crust_type = plot_richness(physeq.prune.rarefy, x="Crust_type", color=("Crust_type"), measures=c("Chao1")) + geom_boxplot() + ggtitle("Mojave Biocrust Bacterial Alpha Diversity (Chao1) by Crust type") + stat_compare_means(method = "anova", label.y = 3000) + stat_compare_means(aes(label=..p.signif..), method = "t.test", ref.group = "LAC", label.y = 2800, hide.ns = TRUE) + theme_bw() + theme(plot.title = element_text(hjust = 0.5))

physeq.prune.rarefy.plot.richness.Crust_type
```

```{r}
physeq.prune.rarefy.ps.plot.richness = plot_richness(physeq.prune.rarefy, x="Crust_type", color=("Crust_type"), measures=c("Observed")) + geom_boxplot() + ggtitle("Mojave Biocrust Fungal Alpha Diversity (Chao1) by Crust type") + stat_compare_means(method = "anova", label.y = 400) + stat_compare_means(aes(label=..p.signif..), method = "t.test", ref.group = "GLC", label.y = 380, hide.ns = TRUE) + theme_bw() + theme(plot.title = element_text(hjust = 0.5))

physeq.prune.rarefy.ps.plot.richness
```

```{r}
pdf("./Figures/Mojave Biocrust Fungal Alpha Diversity (Observed) by Crust type.pdf", width = 8, height = 5)
physeq.prune.rarefy.ps.plot.richness
dev.off()
```


###STEP11.2: Plot Alpha diversity (Observed) by Site and Layer

###Reorder Site
```{r}
#physeq.prune.rarefy.SF = subset_samples(physeq.prune, Layer=="Surface")
#sample_data(physeq.prune.rarefy.SF)$Site = factor(sample_data(physeq.prune.rarefy.SF)$Site, levels = c("JTNP","CIMA", "GMT", "KELSO"))
```

```{r}
sample_data(physeq.prune.rarefy)$Site_layer = factor(sample_data(physeq.prune.rarefy)$Site_layer, levels = c("JTNP_SF","CIMA_SF", "GMT_SF", "KELSO_SF", "JTNP_Sub","CIMA_Sub", "GMT_Sub", "KELSO_Sub"))
```


###ANOVA
```{r}
alpha.diversity = estimate_richness(physeq.prune.rarefy, measures = c("Observed"))
data.anova = cbind(sample_data(physeq.prune.rarefy), alpha.diversity)
physeq.prune.rarefy.anova = aov(Observed ~ Site_layer, data.anova)
summary(physeq.prune.rarefy.anova)
physeq.prune.rarefy.anova.Site.Layer = aov(Observed ~ Site*Layer, data.anova)
summary(physeq.prune.rarefy.anova.Site.Layer)
```

```{r}
tukey.ps = TukeyHSD(x=physeq.prune.rarefy.anova, 'Site_layer', conf.level = 0.95)
```

```{r}
generate_label_df <- function(tukey.ps, variable){
 
     # Extract labels and factor levels from Tukey post-hoc 
     Tukey.levels <- tukey.ps[[variable]][,4]
     Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])
     
     #I need to put the labels in the same order as in the boxplot :
     Tukey.labels$treatment=rownames(Tukey.labels)
     Tukey.labels=Tukey.labels[order(Tukey.labels$treatment) , ]
     return(Tukey.labels)
     }
```


```{r}
LABELS=generate_label_df(tukey.ps, 'Site_layer')
```

```{r}
names(LABELS) = c('Letters','Site_layer')
```

```{r}
ylabel <- data.frame("ylabel" = c(245,280,235,345,215,295,210,320))
```

```{r}
LABELS$ylabel<-ylabel$ylabel
```

```{r}
LABELS
```


```{r warning=FALSE}
physeq.prune.rarefy.ps.plot.richness.site.layer = plot_richness(physeq.prune.rarefy, x="Site_layer", color=("Site_layer"), measures=c("Observed")) + geom_boxplot() + ggtitle("Mojave Biocrust Fungal Alpha Diversity by Site and Layer") + stat_compare_means(method = "anova", label.y = 350) + theme(plot.title = element_text(hjust = 0.5)) + geom_text(data=LABELS, aes(x=LABELS$Site, y = LABELS$ylabel, label = LABELS$Letters)) + theme_bw()

physeq.prune.rarefy.ps.plot.richness.site.layer
```

```{r}
#physeq.prune.rarefy.plot.richness.Site_layer = plot_richness(physeq.prune.rarefy.SF, x="Site", color=("Site"), measures=c("Observed")) + geom_boxplot(lwd=1,fill=c("#999999", "gold2", "cadetblue3", "palegreen2")) + ggtitle("Fungal Alpha Diversity by Site") + stat_compare_means(method = "anova", label.y = 520) + theme(plot.title = element_text(hjust = 0.5)) + scale_colour_manual(values = c("#000000", "#E69F00", "#56B4E9", "#009E73")) + geom_text(data=LABELS, aes(x=LABELS$Site, y = 500, label = LABELS$Letters))

#physeq.prune.rarefy.plot.richness.Site_layer 
```

```{r warning=FALSE}
pdf("./Figures/Mojave Biocrust Fungal Alpha Diversity by Site and Layer.pdf", width = 8, height = 5)
physeq.prune.rarefy.ps.plot.richness.site.layer
dev.off()
```


```{r warning=FALSE}
physeq.prune.rarefy.ps.plot.richness.site.layer = plot_richness(physeq.prune.rarefy, x="Site_layer", color=("Site_layer"), measures=c("Observed")) + geom_boxplot() + ggtitle("Mojave Biocrust Fungal Alpha Diversity (Observed) by Site and Layer") + stat_compare_means(method = "anova", label.y = 400) + stat_compare_means(aes(label=..p.signif..), method = "t.test", ref.group = "GMT_Sub", label.y = 380, hide.ns = TRUE) + theme_bw() + theme(plot.title = element_text(hjust = 0.5))

physeq.prune.rarefy.ps.plot.richness.site.layer
```

To save Alpha diversity (Chao1) by Site plot as a pdf into your current working directory.

```{r warning=FALSE}
pdf("./Figures/Mojave Biocrust Fungal Alpha Diversity (Observed) by Site and Layer.pdf", width = 8, height = 5)
physeq.prune.rarefy.ps.plot.richness.site.layer
dev.off()
```


MS: Figure 2 A (Surface only)

```{r}
physeq.prune.rarefy.SF = subset_samples(physeq.prune.rarefy, Layer=="Surface")
```

```{r}
alpha.diversity.SF = estimate_richness(physeq.prune.rarefy.SF, measures = c("Observed"))
data.anova.SF = cbind(sample_data(physeq.prune.rarefy.SF), alpha.diversity.SF)
physeq.prune.rarefy.anova.SF = aov(Observed ~ Site, data.anova.SF)
summary(physeq.prune.rarefy.anova.SF)
```

```{r}
tukey.ps = TukeyHSD(x=physeq.prune.rarefy.anova.SF, 'Site', conf.level = 0.95)

generate_label_df <- function(tukey.ps, variable){
 
     # Extract labels and factor levels from Tukey post-hoc 
     Tukey.levels <- tukey.ps[[variable]][,4]
     Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])
     
     #I need to put the labels in the same order as in the boxplot :
     Tukey.labels$treatment=rownames(Tukey.labels)
     Tukey.labels=Tukey.labels[order(Tukey.labels$treatment) , ]
     return(Tukey.labels)
     }

LABELS=generate_label_df(tukey.ps, 'Site')
names(LABELS) = c('Letters','Site')
```

```{r}
LABELS
```

```{r}
ylabel <- data.frame("ylabel" = c(245,235,215,210))
LABELS$ylabel<-ylabel$ylabel
LABELS
```

```{r}
physeq.prune.rarefy.plot.richness.Site = plot_richness(physeq.prune.rarefy.SF, x="Site", color=("Site"), measures=c("Observed")) + geom_boxplot(lwd=0.5) + ggtitle("Mojave Biocrust Fungal Alpha Diversity by Site") + stat_compare_means(method = "anova", label.y = 280) + theme(plot.title = element_text(hjust = 0.5)) + geom_text(data=LABELS, aes(x=LABELS$Site, y = LABELS$ylabel , label = LABELS$Letters)) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_colour_brewer(palette="Set1")

physeq.prune.rarefy.plot.richness.Site
```

```{r warning=FALSE}
pdf("./Figures/Fig2A_Fungal_Alpha_Div_by_Site.pdf", width = 6, height = 3.8 )
physeq.prune.rarefy.plot.richness.Site
dev.off()
```

Subsurface

```{r}
physeq.prune.rarefy.SUB = subset_samples(physeq.prune.rarefy, Layer=="Subsurface")
```

```{r}
physeq.prune.rarefy.plot.richness.Site.SUB = plot_richness(physeq.prune.rarefy.SUB, x="Site", color=("Site"), measures=c("Observed")) + geom_boxplot(lwd=1) + ggtitle("Mojave Biocrust Fungal Alpha Diversity by Site") + stat_compare_means(method = "anova", label.y = 280) + theme(plot.title = element_text(hjust = 0.5)) + geom_text(data=LABELS, aes(x=LABELS$Site, y = 270 , label = LABELS$Letters)) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

physeq.prune.rarefy.plot.richness.Site.SUB
```

MS: Figure 2 C

```{r}
alpha.diversity = estimate_richness(physeq.prune.rarefy, measures = c("Observed"))
data.anova = cbind(sample_data(physeq.prune.rarefy), alpha.diversity)
t.test(Observed ~ Layer, data.anova)
```

```{r}
physeq.prune.rarefy.plot.richness.Layer = plot_richness(physeq.prune.rarefy, x="Layer", color=("Layer"), measures=c("Observed")) + geom_boxplot(lwd=0.5) + ggtitle("Mojave Biocrust Fungal Alpha Diversity by Layer") + stat_compare_means(method = "t.test", label.y = 350) + theme(plot.title = element_text(hjust = 0.5)) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_colour_brewer(palette="Set1")

physeq.prune.rarefy.plot.richness.Layer
```

```{r warning=FALSE}
pdf("./Figures/Fig2C_Fungal_Alpha_Div_by_Layer.pdf", width = 6, height = 3.8 )
physeq.prune.rarefy.plot.richness.Layer
dev.off()
```

MS: Figure 2 E Fungal alpha Div by Crust type (Surface only)

```{r}
alpha.diversity = estimate_richness(physeq.prune.rarefy.SF, measures = c("Observed"))
data.anova = cbind(sample_data(physeq.prune.rarefy.SF), alpha.diversity)
physeq.prune.rarefy.anova.Crusttype.SF = aov(Observed ~ Crust_type, data.anova)
summary(physeq.prune.rarefy.anova.Crusttype.SF)
```

```{r}
tukey.ps = TukeyHSD(x=physeq.prune.rarefy.anova.Crusttype.SF, 'Crust_type', conf.level = 0.95)
```

```{r}
generate_label_df <- function(tukey.ps, variable){
 
     # Extract labels and factor levels from Tukey post-hoc 
     Tukey.levels <- tukey.ps[[variable]][,4]
     Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])
     
     #I need to put the labels in the same order as in the boxplot :
     Tukey.labels$treatment=rownames(Tukey.labels)
     Tukey.labels=Tukey.labels[order(Tukey.labels$treatment) , ]
     return(Tukey.labels)
     }
```

```{r}
LABELS=generate_label_df(tukey.ps, 'Crust_type')
```

```{r}
names(LABELS) = c('Letters','Crust_type')
```

```{r}
LABELS
```

```{r}
ylabel <- data.frame("ylabel" = c(215,95,225,235,235))
```

```{r}
LABELS$ylabel<-ylabel$ylabel
```

```{r}
physeq.prune.rarefy.plot.richness.Crust_type = plot_richness(physeq.prune.rarefy.SF, x="Crust_type", color=("Crust_type"), measures=c("Observed")) + geom_boxplot(lwd=0.5) + ggtitle("Mojave Biocrust Fungal Alpha Diversity by Crust type") + stat_compare_means(method = "anova", label.y = 258) + theme(plot.title = element_text(hjust = 0.5)) + geom_text(data=LABELS, aes(x=LABELS$Crust_type, y = LABELS$ylabel , label = LABELS$Letters)) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_colour_brewer(palette="Set1")

physeq.prune.rarefy.plot.richness.Crust_type
```

```{r warning=FALSE}
pdf("./Figures/Fig2E_Fungal_Alpha_Div_by_Crust_type.pdf", width = 6, height = 3.8 )
physeq.prune.rarefy.plot.richness.Crust_type
dev.off()
```

SUBsurface

```{r}
physeq.prune.rarefy.plot.richness.Crust_type.SUB = plot_richness(physeq.prune.rarefy.SUB, x="Crust_type", color=("Crust_type"), measures=c("Observed")) + geom_boxplot(lwd=0.8) + ggtitle("Mojave Biocrust Fungal Alpha Diversity by Crust type") + stat_compare_means(method = "anova", label.y = 358) + theme(plot.title = element_text(hjust = 0.5)) + theme_bw()

physeq.prune.rarefy.plot.richness.Crust_type.SUB
```

###Alpha div by crust type
###Reorder Crust_type
```{r}
#physeq.prune.rarefy.SF = subset_samples(physeq.prune, Layer=="Surface")
#sample_data(physeq.prune.rarefy.SF)$Crust_type = factor(sample_data(physeq.prune.rarefy.SF)$Crust_type, levels = c("LAC","CLC", "GLC", "RMC", "SMC"))
sample_data(physeq.prune.rarefy)$Crust_type = factor(sample_data(physeq.prune.rarefy)$Crust_type, levels = c("LAC","CLC", "GLC", "RMC", "SMC","LAC.Sub","CLC.Sub", "GLC.Sub", "RMC.Sub", "SMC.Sub"))
```

###ANOVA
```{r}
alpha.diversity = estimate_richness(physeq.prune.rarefy, measures = c("Observed"))
data.anova = cbind(sample_data(physeq.prune.rarefy), alpha.diversity)
physeq.prune.rarefy.anova = aov(Observed ~ Crust_type, data.anova)
summary(physeq.prune.rarefy.anova)
physeq.prune.rarefy.anova.Crusttype_layer = aov(Observed ~ Crust_type*Layer, data.anova)
summary(physeq.prune.rarefy.anova.Crusttype_layer)
```

```{r}
tukey.ps = TukeyHSD(x=physeq.prune.rarefy.anova, 'Crust_type', conf.level = 0.95)
```

```{r}
generate_label_df <- function(tukey.ps, variable){
 
     # Extract labels and factor levels from Tukey post-hoc 
     Tukey.levels <- tukey.ps[[variable]][,4]
     Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])
     
     #I need to put the labels in the same order as in the boxplot :
     Tukey.labels$treatment=rownames(Tukey.labels)
     Tukey.labels=Tukey.labels[order(Tukey.labels$treatment) , ]
     return(Tukey.labels)
     }
```


```{r}
LABELS=generate_label_df(tukey.ps, 'Crust_type')
```

```{r}
names(LABELS) = c('Letters','Crust_type')
```

```{r}
ylabel <- data.frame("ylabel" = c(219,320,90,215,225,345,235,300,235,200))
```

```{r}
LABELS$ylabel<-ylabel$ylabel
```

```{r}
#physeq.prune.rarefy.plot.richness.Crust_type = plot_richness(physeq.prune.rarefy.SF, x="Crust_type", color=("Crust_type"), measures=c("Observed")) + geom_boxplot(lwd=1,fill=c("#999999", "gold2", "cadetblue3", "palegreen2", "lightpink")) + ggtitle("Fungal Alpha Diversity by Crust_type") + stat_compare_means(method = "anova", label.y = 520) + theme(plot.title = element_text(hjust = 0.5)) + scale_colour_manual(values = c("#000000", "#E69F00", "#56B4E9", "#009E73", "#CC79A7")) + geom_text(data=LABELS, aes(x=LABELS$Crust_type, y = 500, label = LABELS$Letters))

#physeq.prune.rarefy.plot.richness.Crust_type
```

```{r}
physeq.prune.rarefy.plot.richness.Crust_type = plot_richness(physeq.prune.rarefy, x="Crust_type", color=("Crust_type"), measures=c("Observed")) + geom_boxplot() + ggtitle("Mojave Biocrust Fungall Alpha Diversity by Crust type and Layer") + stat_compare_means(method = "anova", label.y = 380) + theme(plot.title = element_text(hjust = 0.5)) + geom_text(data=LABELS, aes(x=LABELS$Crust_type, y = LABELS$ylabel , label = LABELS$Letters)) + theme_bw()

physeq.prune.rarefy.plot.richness.Crust_type
```

```{r warning=FALSE}
pdf("./Figures/Mojave Biocrust Fungall Alpha Diversity by Crust type and Layer.pdf", width = 8, height = 5)
physeq.prune.rarefy.plot.richness.Crust_type
dev.off()
```

###STEP11.3: Plotting alpha diversity for each selected taxonomy level
Subset fungal OTUs by class

```{r}
FungiClass = as.character(get_taxa_unique(physeq.prune.rarefy.SF, "Class"))
FungiClass = FungiClass[complete.cases(FungiClass)]
```

Save pdf of plot_richness results for each fungal class in the loop which go through FungiClass
Note: following code plot alpha diversity by Site and Layer
```{r}
pdf("./Figures/Fungal Class Alpha Diversity Observed by Site PLPA250.pdf", width = 8, height = 5 )
for (i in FungiClass) {
  APS = subset_taxa(physeq.prune.rarefy.SF, Class == i)
  APS.plot = plot_richness(APS, x="Site", color =("Site"), measures=c("Observed")) + geom_boxplot(lwd=1,fill=c("#999999", "gold2", "cadetblue3", "palegreen2")) + ggtitle(i) + stat_compare_means(method = "anova") + theme(plot.title = element_text(hjust = 0.5)) + scale_colour_manual(values = c("#000000", "#E69F00", "#56B4E9", "#009E73"))
  print(APS.plot)
}
dev.off()
```

Save pdf of plot_richness results for each fungal class in the loop which go through FungiClass
Note: following code plot alpha diversity by Crust type

```{r}
pdf("./Figures/Fungal Class Alpha Diversity (Observed) by Crust type.pdf", width = 8, height = 5 )
for (i in FungiClass) {
  APS = subset_taxa(physeq.prune.rarefy, Class == i)
  APS.plot = plot_richness(APS, x="Crust_type", color =("Crust_type"), measures=c("Observed")) + geom_boxplot() + ggtitle(i) + stat_compare_means(method = "anova") + theme(plot.title = element_text(hjust = 0.5))
  print(APS.plot)
}
dev.off()
```


MS: Figure S1A comparing taxonmic differences by Site

```{r}
physeq.prune.rarefy.SF = subset_samples(physeq.prune.rarefy, Layer=="Surface")
```

```{r}
FungiClass = as.character(get_taxa_unique(physeq.prune.rarefy.SF, "Class"))
FungiClass = FungiClass[complete.cases(FungiClass)]
```

```{r}
pdf("./Figures/Fungal_Class_Alpha_(Observed)_Diversity_by_Site_surface_only.pdf", width = 6, height = 3.8)
for (i in FungiClass) {
  APS = subset_taxa(physeq.prune.rarefy.SF, Class == i)
  APS.plot = plot_richness(APS, x="Site", color =("Site"), measures=c("Observed")) + geom_boxplot(lwd=0.5) + ggtitle(i) + stat_compare_means(method = "anova") + theme(plot.title = element_text(hjust = 0.5)) + theme_bw() + scale_colour_brewer(palette="Set1") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
  print(APS.plot)
}
dev.off()
```

MS: Figure S1B comparing taxonmic differences by Layer

```{r}
FungiClass.all = as.character(get_taxa_unique(physeq.prune.rarefy, "Class"))
FungiClass.all = FungiClass.all[complete.cases(FungiClass.all)]
FungiClass.all
```

```{r}
pdf("./Figures/Bacterial_Phylum_Alpha_(Observed)_Diversity_by_Layer.pdf", width = 6, height = 3.8)
for (i in FungiClass.all) {
  APS = subset_taxa(physeq.prune.rarefy, Class == i)
  APS.plot = plot_richness(APS, x="Layer", color =("Layer"), measures=c("Observed")) + geom_boxplot(lwd=0.5) + ggtitle(i) + stat_compare_means(method = "t.test", label.x.npc = c("right")) + theme(plot.title = element_text(hjust = 0.5)) + theme_bw() + scale_colour_brewer(palette="Set1") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
  print(APS.plot)
}
dev.off()
```

```{r}
APS = subset_taxa(physeq.prune.rarefy, Class == "Sordariomycetes")
APS.plot = plot_richness(APS, x="Layer", color =("Layer"), measures=c("Observed")) + geom_boxplot(lwd=0.5) + ggtitle("Sordariomycetes") + stat_compare_means(method = "t.test", label.x.npc = c("right")) + theme(plot.title = element_text(hjust = 0.5)) + theme_bw() + scale_colour_brewer(palette="Set1") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
print(APS.plot)
```

```{r}
APS = subset_taxa(physeq.prune.rarefy, Class == "Sordariomycetes")
alpha.diversity.Sordariomycetes = estimate_richness(APS, measures = c("Observed"))
data.anova.Sordariomycetes = cbind(sample_data(APS), alpha.diversity.Sordariomycetes)
t.test(Observed ~ Layer, data.anova.Sordariomycetes)
```

```{r}
APS = subset_taxa(physeq.prune.rarefy, Class == "Schizosaccharomycetes")
alpha.diversity.Sordariomycetes = estimate_richness(APS, measures = c("Observed"))
data.anova.Sordariomycetes = cbind(sample_data(APS), alpha.diversity.Sordariomycetes)
t.test(Observed ~ Layer, data.anova.Sordariomycetes)
```


MS: Figure S1C comparing taxonmic differences by Crust types

```{r}
pdf("./Figures/Bacterial_Phylum_Alpha_(Observed)_Diversity_by_Crust_type_surface_only.pdf", width = 6, height = 3.8 )
for (i in FungiClass) {
  APS = subset_taxa(physeq.prune.rarefy.SF, Class == i)
  APS.plot = plot_richness(APS, x="Crust_type", color =("Crust_type"), measures=c("Observed")) + geom_boxplot(lwd=0.5) + ggtitle(i) + stat_compare_means(method = "anova") + theme(plot.title = element_text(hjust = 0.5)) + theme_bw() + scale_colour_brewer(palette="Set1") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
  print(APS.plot)
}
dev.off()
```


###STEP11.4 Summarize taxa composition

```{r}
taxa.summary.by.phylum = summarize_taxa(physeq.prune.rarefy, "Phylum")
```

######To save taxa summary into text file under your current working directory.

```{r warning=FALSE}
write.table(taxa.summary.by.phylum, file = "taxa.summary.by.phylum.txt", sep ="\t")
```


###STEP12.1 Taxonomic composition

MS: Figure 3A


```{r}
physeq.prune.rarefy
```


```{r}
physeq.prune.rarefy = subset_taxa(physeq.prune.rarefy, Kingdom == "Fungi")
physeq.prune.rarefy
```


```{r fig.height=5, fig.width=8, fig.align="center"}
tax_layer = plot_bar(ps.glom.class, x="Layer", fill = "Class") + 
  geom_bar(aes(color=Class, fill=Class), stat="identity", position="stack") +
  theme_bw() 
```

```{r fig.height=5, fig.width=8, fig.align="center"}
tax_layer
```

```{r}
ps.glom.class = tax_glom(physeq.prune.rarefy, taxrank = "Class")
```

```{r fig.height=5, fig.width=8, fig.align="center"}
taxcom_layer = ggplot(data = psmelt(physeq.prune.rarefy), mapping = aes_string(x = "Layer" ,y = "Abundance", fill = "Class" )) + 
  geom_bar(stat="identity", position="fill") + 
  ggtitle("Fungal Taxonomic Composition (Class level) by Layer")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
print(taxcom_layer)
```

```{r}
png("./Figures/Fig3A_Fungal_Taxonomic_Composition_(Class_level)_by_Layer.png", units="in", width = 8, height = 5.8, res = 300 )
taxcom_layer
dev.off()
```

```{r fig.height=5, fig.width=8, fig.align="center"}
pdf("./Figures/Fig3A_Fungal_Taxonomic_Composition_(Class_level)_by_Layer.pdf", width = 8, height = 5.8 )
taxcom_layer
dev.off()
```

MS: Figure S1A

```{r fig.height=5, fig.width=8, fig.align="center"}
taxcom_site = ggplot(data = psmelt(physeq.prune.rarefy.SF), mapping = aes_string(x = "Site" ,y = "Abundance", fill = "Class" )) + 
  geom_bar(stat="identity", position="fill") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_bw()
print(taxcom_site)
```

```{r fig.height=5, fig.width=8, fig.align="center"}
pdf("./Figures/FigS1A_Fungal_Taxonomic_Composition_(Class_level)_by_Site.pdf", width = 8, height = 5.8 )
taxcom_site
dev.off()
```



MS: Figure S1B

```{r fig.height=5, fig.width=8, fig.align="center"}
taxcom_crust_type = ggplot(data = psmelt(physeq.prune.rarefy.SF), mapping = aes_string(x = "Crust_type" ,y = "Abundance", fill = "Class" )) + 
  geom_bar(stat="identity", position="fill") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_bw()
print(taxcom_crust_type)
```

```{r fig.height=5, fig.width=8, fig.align="center"}
pdf("./Figures/FigS1B_Fungal_Taxonomic_Composition_(Class_level)_by_Crust_type.pdf", width = 8, height = 5.8 )
taxcom_crust_type
dev.off()
```






```{r fig.height=5, fig.width=8, fig.align="center"}
taxcom_sitelayer = ggplot(data = psmelt(physeq.prune.rarefy), mapping = aes_string(x = "Site_layer" ,y = "Abundance", fill = "Class" )) + 
  geom_bar(stat="identity", position="fill") + 
  ggtitle("Fungal Taxonomic Composition (Class level) by Site and Layer")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
#print(taxcom_sitelayer)
```

```{r fig.height=5, fig.width=8, fig.align="center"}
pdf("./Figures/Fungal Taxonomic Composition (Class level) by Site and Layer.pdf", width = 12, height = 8 )
taxcom_sitelayer
dev.off()
```

```{r fig.height=5, fig.width=8, fig.align="center"}
taxcom_crusttype = ggplot(data = psmelt(physeq.prune.rarefy), mapping = aes_string(x = "Crust_type" ,y = "Abundance", fill = "Class" )) + 
  geom_bar(stat="identity", position="fill") + 
  ggtitle("Fungal Taxonomic Composition (Class level) by Crust type")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r fig.height=5, fig.width=8, fig.align="center"}
pdf("./Figures/Fungal Taxonomic Composition (Class level) by Crust type.pdf", width = 12, height = 8 )
taxcom_crusttype
dev.off()
```


```{r fig.height=5, fig.width=8, fig.align="center"}
taxcom_site = ggplot(data = psmelt(physeq.prune.rarefy.SF), mapping = aes_string(x = "Site" ,y = "Abundance", fill = "Class" )) + 
  geom_bar(stat="identity", position="fill") + 
  ggtitle("Fungal Taxonomic Composition by Site")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(plot.title = element_text(hjust = 0.5))

#print(taxcom_sitelayer)
```

```{r fig.height=5, fig.width=8, fig.align="center"}
pdf("./Figures/Fungal Taxonomic Composition by Site.pdf", width = 10, height = 6 )
taxcom_site
dev.off()
```



###Venn

```{r}
physeq.prune.rarefy = subset_taxa(physeq.prune.rarefy, Kingdom != "Rhizaria")
physeq.prune.rarefy = subset_taxa(physeq.prune.rarefy, Kingdom != "Chromista")
physeq.prune.rarefy = subset_taxa(physeq.prune.rarefy, Kingdom != "Unassigned")
physeq.prune.rarefy = subset_taxa(physeq.prune.rarefy, Kingdom != "D_0__Eukaryota")
physeq.prune.rarefy = subset_taxa(physeq.prune.rarefy, Kingdom != "NA")
physeq.prune.rarefy
```

###Subset for Surface and Subsurface
```{r}
physeq.prune.rarefy.SF = subset_samples(physeq.prune.rarefy, Layer=="Surface")
physeq.prune.rarefy.SF = prune_taxa(taxa_sums(physeq.prune.rarefy.SF) > 1, physeq.prune.rarefy.SF)
physeq.prune.rarefy.SF
```

```{r}
physeq.prune.rarefy.SUB = subset_samples(physeq.prune.rarefy, Layer=="Subsurface")
physeq.prune.rarefy.SUB = prune_taxa(taxa_sums(physeq.prune.rarefy.SUB) > 1, physeq.prune.rarefy.SUB)
physeq.prune.rarefy.SUB
```

```{r}
OTU.name.physeq.prune.rarefy.SF = rownames((otu_table(physeq.prune.rarefy.SF)))
```

```{r}
OTU.name.physeq.prune.rarefy.SUB = rownames((otu_table(physeq.prune.rarefy.SUB)))
```

```{r}
## plot venn diagrams
      
# Surface vs Subsurface
pdf("./Figures/Fungal_Surface_Subsurface_venn.pdf", width = 8, height = 8)
ven_test = venn_diagram2(OTU.name.physeq.prune.rarefy.SF, OTU.name.physeq.prune.rarefy.SUB, "Surface", "Subsurface", colors= c("#009E73", "#CC79A7"), euler=TRUE, main = "Fungal Surface-Subsurface Venn")
dev.off()
```



##Lichen

```{r}
physeq.prune.rarefy.Lecanoromycetes = subset_taxa(physeq.prune.rarefy, Class == "Lecanoromycetes")
physeq.prune.rarefy.Lecanoromycetes
```

```{r}
plot_bar(physeq.prune.rarefy.Lecanoromycetes, x = "Crust_type", fill = "Order")
```


```{r fig.height=5, fig.width=8, fig.align="center"}
taxcom_Lecanoromycetes = ggplot(data = psmelt(physeq.prune.rarefy.Lecanoromycetes), mapping = aes_string(x = "Crust_type" ,
                                                                                                         y = "Abundance", fill = "Order" )) + 
  geom_bar(stat="identity", position="fill") + 
  ggtitle("Lecanoromycetes Taxonomic Composition (Order level) by Crust_type")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
print(taxcom_Lecanoromycetes)
```

```{r}
pdf("./Figures/taxcom_Lecanoromycetes.pdf", width = 8, height = 5)
taxcom_Lecanoromycetes
dev.off()

```






