---
title: "Mojave_BacFun_network"
author: "Nat Pombubpa"
date: "5/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(circlize)
```


## Mojave Bacterial-Fungal Network (after SpiecEasi step)
##CLC.SF
```{r}
cdf = read.table("EasiFunBacNet.SF.CLC.top100.e2L1000.csv", header=T,sep=",")
head(cdf)
```

```{r}
Bac.net.taxmat <- read.table("BAC.network.taxonomy.csv", header=T,sep=",")
Fun.net.taxmat <- read.table("FG.network.taxonomy.csv", header=T,sep=",")
tax.circlize = rbind(Fun.net.taxmat,Bac.net.taxmat)
head(tax.circlize)
```

```{r}
cdf.w.tax = data.frame(OTU1=tax.circlize[match(cdf$OTU1, tax.circlize$OTU.ID), 4], OTU2=tax.circlize[match(cdf$OTU2, tax.circlize$OTU.ID), 4], WEIGHT=cdf$WEIGHT)
cdf.w.tax
```

```{r}
cdf.w.tax$OTU1 <- as.character(cdf.w.tax$OTU1)
cdf.w.tax$OTU2 <- as.character(cdf.w.tax$OTU2)
str(cdf.w.tax)
```

```{r}
cdf.w.tax[is.na(cdf.w.tax)] <- "unidentified"
```

clear any object that might be drawn by circlize before.

```{r}
circos.clear()
```

Plot network from "cdf" data frame that was constructed from Phyloseq plot network

link.visible: plot any of the collumn#3 that similarity is greater than certain number in this case 0.8.
for (si in get.all.setor.index()) is used to re-label OTUs name, so they are not on top of each others. (You can try plotting without for (si in get.all.setor.index()) to see how it would look )

```{r}
chordDiagram(cdf.w.tax, annotationTrack = "grid", annotationTrackHeight = 0.01, preAllocateTracks = 1, link.visible = cdf.w.tax[[3]] > 0.85) 
for(si in get.all.sector.index()) {
    xlim = get.cell.meta.data("xlim", sector.index = si, track.index = 1)
    ylim = get.cell.meta.data("ylim", sector.index = si, track.index = 1)
    circos.text(mean(xlim), ylim[1], si, facing = "clockwise", adj = c(-0.1, -0.2),
    niceFacing = TRUE, cex = 1.2, col = "black", sector.index = si, track.index = 1)
} 

circos.clear()
```

```{r}
pdf("Figures/Biocrust Fungal-Bacteria Community Network Analysis (CLC Surface top 100 OTUs-cross only).pdf", width = 18, height = 18)
chordDiagram(cdf.w.tax, annotationTrack = "grid", annotationTrackHeight = 0.01, preAllocateTracks = 1, link.visible = cdf.w.tax[[3]] > 0.85) 
for(si in get.all.sector.index()) {
    xlim = get.cell.meta.data("xlim", sector.index = si, track.index = 1)
    ylim = get.cell.meta.data("ylim", sector.index = si, track.index = 1)
    circos.text(mean(xlim), ylim[1], si, facing = "clockwise", adj = c(-0.1, -0.2),
    niceFacing = TRUE, cex = 1.2, col = "black", sector.index = si, track.index = 1)
} 

title(main = "Biocrust Fungal-Bacterial Community Network Analysis (CLC surface top 100 OTUs-cross domain only)")
circos.clear()
dev.off()
```


