---
title: "Mojave_Fun_Indic_Species"
author: "Nat Pombubpa"
date: "Updated on May 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Documentation for Qiime2 Silva extracted length classifier on Sheephole crusts samples
This is a documentation for otu table, taxonomy table, and tree which were generated from Qiime2 using DADA2 and Silva database classifier extracted lenght (515-806). 

###STEP1: Load all necessary packages for analysis
More information about Phyloseq can be found at the following link: [Phyloseq](https://joey711.github.io/phyloseq/)
If you get error in this step, you probably need to install any packages which causes error.

```{r warning=FALSE, message=FALSE}
library(ape)
library(vegan)
library(plyr)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(data.table)
library(tidyr)
library(labdsv)
```

```{r}
meta = read.table("Mojave_mappingfile_8-Aug-2018.txt",header=TRUE,row.names=1,sep="\t",stringsAsFactors=FALSE)
```

If need to keep or remove some samples
```{r}
meta <- meta[which(meta$Layer %in% c("Surface")),]
```

```{r}
sampleData <- sample_data(meta)
```

###STEP5: Import OTU table

```{r}
FG <- read.table("FGoriginalcopy.txt",header=T,sep="\t",row.names=1)
FGotus <- dplyr::select(FG, -(Taxonomy:Citation.Source))
FGotumat <- as(as.matrix(FGotus), "matrix")
FGOTU = otu_table(FGotumat, taxa_are_rows = TRUE)
```

```{r}
FG
```


Check imported OTU table

```{r warning=FALSE}
head(FGOTU)
```

```{r}
colnames(FG)
```


```{r}
FGtaxmat <- dplyr::select(FG, OTU.ID, Taxonomy, Confidence.Ranking, Trophic.Mode, Guild, Growth.Morphology)
FGtaxmat <- as(as.matrix(FGtaxmat),"matrix")
FGTAX = tax_table(FGtaxmat)
```

###STEP7: Construct Phyloseq object
To construct phyloseq object, otu table, taxonomy table, and sampleData are required. Phylogenetic tree can be included, but it is not necessary for constructing phyloseq object.
Construct Phyloseq object called "Physeq"

```{r warning=FALSE}
physeq = phyloseq(FGOTU,FGTAX,sampleData)
```

###STEP8: Check phyloseq object
This should indicate that your physeq is a "phyloseq-class experiment-level object"

```{r}
physeq
```

###STEP9: Remove singletons
Remove any OTUs that present only one time.

```{r}
physeq.prune = prune_taxa(taxa_sums(physeq) > 1, physeq)
```

###STEP10: Plot read counts to check dataset
Check read counts: any samples that have very low reads should be removed.
[Ref](http://evomics.org/wp-content/uploads/2016/01/phyloseq-Lab-01-Answers.html)

```{r}
readcount = data.table(as(sample_data(physeq.prune), "data.frame"),
                 TotalReads = sample_sums(physeq.prune), 
                 keep.rownames = TRUE)
setnames(readcount, "rn", "SampleID")

#For plotting, use command below.
#SeqDepth = ggplot(readcount, aes(TotalReads)) + geom_histogram() + ggtitle("Sequencing Depth")
```

TotalReads of all the samples can be in this table (select only SampleID and TotalReads columns).
In order to check samples with low number of reads, "order()" can be used to sort "TotalReads" column.
In this dataset, N55.Rhizo has very low number of reads, so will will filter this sample out using the next minimum number of reads.
```{r}
readcount = readcount[order(readcount$TotalReads), c("SampleID", "TotalReads")]
```

```{r}
head(readcount)
```

###STEP11: Rarefy OTUs to a minimum number of reads
Rarefy OTUs (remove any samples that has very low number of reads)
If this step work fine, you have successfully imported data to R and completely generate phyloseq object as "physeq.prune.rarefy".
```{r warning=FALSE}
set.seed(711)
physeq.prune.rarefy = rarefy_even_depth(physeq.prune, sample.size = 10497, replace = FALSE, trimOTUs = TRUE)
physeq.prune.rarefy
```

### Indicator Species (Andrew)

Convert phyloseq OTU table to reflect species

Glomerate at Species level (However, not all of species will be unique)
See for ref: https://github.com/joey711/phyloseq/issues/927

```{r}
phyloGlom = physeq.prune.rarefy
```

Subset samples for specific phyla
```{r}
#phyloGlom = subset_taxa(phyloGlom, Phylum == "D_1__Cyanobacteria")
```

```{r}
tax_table(phyloGlom)
```


Instead of using unique Species name, OTU.ID will be used instead becasue they are unique
```{r}
glomTax = tax_table(phyloGlom)[,"OTU.ID"]
```

Combine OTU table and Tax table
```{r}
glomOTU = otu_table(phyloGlom)
glomTable = merge(glomOTU,glomTax,by=0,all=TRUE)
head(glomTable)
```

Use OTU.ID as rowname and remove prior rowname and OTU.ID
```{r}
rownames(glomTable) = glomTable[,"OTU.ID"]
glomTable$Row.names = NULL
glomTable$OTU.ID = NULL
```

```{r}
head(glomTable)
```


Transpose Rows and columns for desired labdsv format
```{r}
ASV = t(glomTable)
```

Remove columns with 0 species counts

```{r}
ASV.no0 = ASV[, colSums(ASV != 0) > 0]
```

```{r}
ASV.no0
```

Subset sampleData phyloseq object for the desired variable (e.g. treatment, depth, crust type, etc.)

```{r}
Crust_type.Mojave = subset(sampleData,select = "Site")
```

```{r}
class(sampleData$Site)
```


Make the varible numeric

```{r}
sampleData$Site
```

```{r}
Crust_type.Mojave$Site = as.numeric(as.factor(Crust_type.Mojave$Site))
Crust_type.Mojave$Site
```


Run indval code from lab.dsv 

```{r}
Crust_type.Mojave.indic = indval(ASV.no0,Crust_type.Mojave$Site)
```

```{r}
summary(Crust_type.Mojave.indic, p = 0.05, type = "short")
```

```{r}
summary(Crust_type.Mojave.indic, p = 0.05, type = "long")
```

Get indval value
```{r}
Crust_type.Mojave.indic$indval
```

```{r}
Crust_type.Mojave.indic$indcls
```

###Get Significant Indicators (203 ASVs)

```{r}
gr <- Crust_type.Mojave.indic$maxcls[Crust_type.Mojave.indic$pval<=0.05]
iv <- Crust_type.Mojave.indic$indcls[Crust_type.Mojave.indic$pval<=0.05]
pv <- Crust_type.Mojave.indic$pval[Crust_type.Mojave.indic$pval<=0.05]
fr <- apply(ASV.no0[,-1]>0, 2, sum)[Crust_type.Mojave.indic$pval<=0.05]
indvalsummary <- data.frame(group=gr, indval=iv, pvalue=pv, freq=fr)
indvalsummary <- indvalsummary[order(indvalsummary$group, -indvalsummary$indval),]
```

```{r}
indvalsummary
```


```{r}
Crust_type.Mojave.indic.indval = as.data.frame(Crust_type.Mojave.indic$indval)
Crust_type.Mojave.indic.indval
```

Match rowname of indval table to summary table and take only 203 matches indicator species

```{r}
Crust_type.Mojave.indic.indval_filtered = Crust_type.Mojave.indic.indval[row.names(Crust_type.Mojave.indic.indval) %in% row.names(indvalsummary),]
Crust_type.Mojave.indic.indval_filtered
```

Rename columns ###Stop here May 23, 10:45
```{r}
names(Crust_type.Mojave.indic.indval_filtered)[names(Crust_type.Mojave.indic.indval_filtered) == "1"] <- "CIMA"
names(Crust_type.Mojave.indic.indval_filtered)[names(Crust_type.Mojave.indic.indval_filtered) == "2"] <- "GMT"
names(Crust_type.Mojave.indic.indval_filtered)[names(Crust_type.Mojave.indic.indval_filtered) == "3"] <- "JTNP"
names(Crust_type.Mojave.indic.indval_filtered)[names(Crust_type.Mojave.indic.indval_filtered) == "4"] <- "KELSO"
#names(Crust_type.Mojave.indic.indval_filtered)[names(Crust_type.Mojave.indic.indval_filtered) == "5"] <- "SMC"
head(Crust_type.Mojave.indic.indval_filtered)
```

Add otu table from filtered indval table
```{r}
indic.otumat <- as(as.matrix(Crust_type.Mojave.indic.indval_filtered), "matrix")
indic.OTU = otu_table(indic.otumat, taxa_are_rows = TRUE)
```

```{r}
head(indic.OTU)
```

add metadata table
```{r warning=FALSE}
indic.meta = read.table("Fun.indic.meta.site.tsv",
                  header=TRUE,row.names=1,
                  sep="\t",stringsAsFactors=FALSE)
```

```{r}
indic.meta
```


```{r warning=FALSE}
indic.sampleData <- sample_data(indic.meta)
```


```{r warning=FALSE}
indic.physeq = phyloseq(indic.OTU,FGTAX,indic.sampleData)
```

```{r}
indic.physeq
```



```{r}
#Heatmap.indic.physeq = prune_taxa(names(sort(taxa_sums(indic.physeq),TRUE)[1:203]),indic.physeq)
```

```{r}
sample_data(indic.physeq)$Crust_type = factor(sample_data(indic.physeq)$Crust_type, levels = c("LAC", "CLC", "GLC", "RMC", "SMC"))
```

```{r}
indic.physeq.TAX = tax_table(indic.physeq)
indic.physeq.OTU.Tax.tab = merge(indvalsummary,indic.physeq.TAX,by=0,all=TRUE)
indic.physeq.OTU.Tax.tab
```

```{r}
write.csv(indic.physeq.OTU.Tax.tab, file="FunGuilds.indvalsummary.site.Mojave.csv")
```

```{r}
Heatmap.indic.physeq = prune_taxa(names(sort(taxa_sums(indic.physeq),TRUE)[1:65]),indic.physeq)
```

```{r}
Heatmap.indic.physeq.plot = plot_heatmap(Heatmap.indic.physeq, "PCoA", "bray", sample.label = NULL ,"Genus", sample.order = c("LAC", "CLC", "GLC", "RMC", "SMC"))+ ggtitle("Bacterial indicator species heatmap (P < 0.05)") + theme(plot.title = element_text(hjust = 0.5))
print(Heatmap.indic.physeq.plot)
```

```{r}
ggsave("./Figures/Heatmap_indicator_species_Genuslevel_0.01P.png", plot = Heatmap.indic.physeq.plot, width = 12, height = 8)
```



###Checking indicator species using abundance table

```{r}
indic.TAX = tax_table(indic.physeq)
```

```{r warning=FALSE}
otu <- read.table("MojaveBacQ2FWLE.otu_table.fix.txt",header=T,sep="\t")
otus.filtered = otu[otu$OTU.ID %in% OTU_SF_merge_crust_type_endemic_df$OTUendemic,]
indic.otus.filtered = otus.filtered[otus.filtered$OTU.ID %in% row.names(indvalsummary),]
indic.otus.filtered.rn = indic.otus.filtered[,-1]
rownames(indic.otus.filtered.rn) = indic.otus.filtered[,1]
indic.otumat.endemic <- as(as.matrix(indic.otus.filtered.rn), "matrix")
indic.OTU.endemic = otu_table(indic.otumat.endemic, taxa_are_rows = TRUE)
```

```{r warning=FALSE}
abundance.indic.physeq = phyloseq(indic.OTU.endemic,indic.TAX,sampleData)
```

```{r}
abundance.indic.physeq
```

```{r}
abundance.indic.physeq.noGLC = subset_samples(abundance.indic.physeq, Crust_type !="GLC")
```

```{r}
Heatmap.abundance.indic.physeq.noGLC.plot = plot_heatmap(abundance.indic.physeq.noGLC, "NMDS", "bray", "Crust_type", "Genus", sample.order = "Crust_type")+ ggtitle("Bacterial endemic indicator spcecies abundance") + theme(plot.title = element_text(hjust = 0.5))
print(Heatmap.abundance.indic.physeq.noGLC.plot)

ggsave("./Figures/Bacterial_endemic_indicator_species_heatmap_Genus_May14.png", plot = Heatmap.abundance.indic.physeq.noGLC.plot, width = 12, height = 8)
```













